\documentclass[draft,final]{vutinfth}

\input{Commands/commands}
\input{Commands/commands_components}
\input{Commands/commands_diagram}
%\input{Commands/commands_listing}



% Define convenience functions to use the author name and the thesis title in the PDF document properties.
\newcommand{\authorname}{Janos Tapolczai} % The author name without titles.
\newcommand{\thesistitle}{Title of the Thesis} % The title of the thesis. The English version should be used, if it exists.

% Set PDF document properties	
\hypersetup{
	pdfpagelayout   = TwoPageRight,           % How the document is shown in PDF viewers (optional).
	linkbordercolor = {Melon},                % The color of the borders of boxes around crosslinks (optional).
	pdfauthor       = {\authorname},          % The author's name in the document properties (optional).
	pdftitle        = {\thesistitle},         % The document's title in the document properties (optional).
	pdfsubject      = {Subject},              % The document's subject in the document properties (optional).
	pdfkeywords     = {a, list, of, keywords} % The document's keywords in the document properties (optional).
}

\setsecnumdepth{subsection} % Enumerate subsections.

\nonzeroparskip             % Create space between paragraphs (optional).
\setlength{\parindent}{0pt} % Remove paragraph identation (optional).

\makeindex      % Use an optional index.
% \makeglossaries % Use an optional glossary.
%\glstocfalse   % Remove the glossaries from the table of contents.

% Set persons with 4 arguments:
%  {title before name}{name}{title after name}{gender}
%  where both titles are optional (i.e. can be given as empty brackets {}).
\setauthor{BSc.}{\authorname}{}{male}
\setadvisor{a.o. Univ.-Prof. Dr.}{Hans Tompits}{}{male}

% For bachelor and master theses:
\setfirstassistant{Pretitle}{Forename Surname}{Posttitle}{male}
\setsecondassistant{Pretitle}{Forename Surname}{Posttitle}{male}
\setthirdassistant{Pretitle}{Forename Surname}{Posttitle}{male}

% Required data.
\setaddress{Address}
\setregnumber{0825077}
\setdate{01}{01}{2016} % Set date with 3 arguments: {day}{month}{year}.
\settitle{\thesistitle}{Design and Implementation of an Agent Architecture combining Emotions and Reasoning}
% \setsubtitle{Optional Subtitle of the Thesis}{Optionaler Untertitel der Arbeit} % Sets English and German version of the subtitle (both can be English or German).

% Select the thesis type: bachelor / master / doctor / phd-school.
% Bachelor:
\setthesis{master}
\setmasterdegree{dipl.} % dipl. / rer.nat. / rer.soc.oec. / master

% For bachelor and master:
\setcurriculum{Computational Intelligence}{Computational Intelligence} % Sets the English and German name of the curriculum.

\begin{document}
	
	\frontmatter % Switches to roman numbering.
	% The structure of the thesis has to conform to
	%  http://www.informatik.tuwien.ac.at/dekanat
	
	\addtitlepage{naustrian} % German title page (not for dissertations at the PhD School).
	% \addtitlepage{english} % English title page.
	\addstatementpage
	
	\begin{acknowledgements*}
		This thesis is the result of two years of work.
		
		
		I would like to thank my mother for tirelessly pursuing my education from my early years on, instilling in me the value of bettering oneself. Without her tremendous efforts, I would doubtlessly not be where I am today.
		
		matthias \"{O}sterreicher
		I would also like my good friends J\"{u}rgen Cito, Manuel Mertl, Bernd Rathmanner, Gerald Schermann for their support and friendship through these years.
		
		Furthermore, I would like to sincerely thank my professors Hans Tompits and Rudolf Freund for their advice, and their forbearance in discussing my admittedly idiosyncratic ideas at length.
	\end{acknowledgements*}
	
	\begin{abstract}
		We design and implement a hybrid AI that combines emotional reactions to stimuli with reasoning about the consequences of actions as a means of developing effective behaviour in a toy world. The fundamental task of our agents is to survive by periodically finding and eating food and by dealing with hostile predators, either through flight or through flight. Agents are free to cooperate, antagonize, or ignore each other and agents with different emotional profiles will pursue different strategies.
		
		The thesis begins with theoretical investigations of different models of computations and their relation to the biological brain. Our proposition is that the brain's function is, to a degree, analogous to a collection of white boxes, observable to each other. Accordingly, our agents are modeled as a collection of loosely coupled components which communicate with each other through messages. Any component is free to read any message and components have no knowledge which others will read messages they insert into the agent's message space.
		
		A relatively simple affective evaluation forms the basis of each agent's decision-making, though it is complemented by a belief generator which makes inferences about future world-states resulting from certain choices. Though evaluation of these future states as well, agents can optimize their behavior, as they can foresee likely positive or negative consequences of their actions.
		
		The thesis ends will a qualitative and a quantitative evaluation. We qualitatively evaluate agents qualitatively by placing agents in a number of simple scenarios and observing whether they perform tasks like collecting food or avoiding predators. We then evaluate them quantitatively by placing various populations into larger scenarios and recording the survival of different personality types over time.
	\end{abstract}
	
	\begin{kurzfassung}
		Wir entwerfen und implementieren eine hybride AI, welche emotionale Reaktionen auf Stimuli mit dem Schließen über den Folgen von Aktionen zum Zwecke der Ausarbeitung effektiven Verhaltens in einer einfachen Welt kombiniert. Die fundamentale Aufgabe unserer Agenten ist es, durch das Finden und Essen von Nahrung, sowie der Handhabung von Raubtieren durch Kampf oder Flucht, zu überleben.
		
		Die Arbeit beginnt mit einer theoretischen Untersuchung mehrerer komputationaler Modelle und ihrem Verhältnus zum biologischen Gehirn. Unserer Behauptung ist, dass das Gehirn zu einem gewissen Grad analog zu einer Ansammlung von \emph{White Boxes} ist, die die Arbeit der jeweils anderen beobachten können. Demgemäß sind unsere Agenten als lose gekoppelte Komponenten modelliert, die über Nachrichten miteinander kommunizieren. Jede Komponente kann jede Nachricht auslesen und keine Komponente weiß, welche anderen die Nachrichten lesen werden, welche sie in den zentralen \emph{Message Space} des Agenten einfügt.
		
		Eine relativ einfache affektive Evaluation ist die Grundlage der Entscheidungsfindung der Agenten, sie wird aber ergänzt durch einen \emph{Belief Generator}, der zukünftige Zustände der Welt als Folge bestimmter Aktionen simuliert. Durch die Evaluation dieser zukünftigen Zustände können Agenten ihr Verhalten anpassen und sowohl die negativen als auch die positiven Folgen ihrer Aktionen absehen.
		
		Die Arbeit endet mit einer qualitativen sowie einer quantitativen Evaluierung. Wir evaluieren die Agenten qualitativ, indem wir sie in einer reihe simpler Beispielszenarien ausprobieren und testen, ob sie einfache Aufgaben, wie das Sammeln von Essen oder die Flucht von Feinden, ausführen. Danach evaluieren wir sie quantativ mittels verschiedener Populationen in eine größeren Szenario, in dem wir die langfristige Überlebensrate verschiedener Persönlichkeitstypen festhalten.
	\end{kurzfassung}
	
	% Select the language of the thesis, e.g., english or naustrian.
	\selectlanguage{english}
	
	% Add a table of contents (toc).
	\tableofcontents % Starred version, i.e., \tableofcontents*, removes the self-entry.
	
	% Switch to arabic numbering and start the enumeration of chapters in the table of content.
	\mainmatter
	
	\chapter{Introduction}
	
	\input{1introduction}
	
	\chapter{Preliminary Considerations}\label{ch:preliminaryConsiderations}
	
	\input{2preliminaryConsiderations}
	
	\chapter{Overall Component Model}\label{ch:componentModel}
	
	\input{3componentModel}
	
	\chapter{Affective Architecture}\label{ch:affectiveArchitecture}.
	
	\input{4affectiveArchitecture}
	
	\chapter{Implementation}\label{ch:implementation}
	
	\input{5implementation}
	
	\chapter{Conclusion}\label{ch:conclusion}
	
	\input{6conclusion}
	
	
	
	% Remove following line for the final thesis.
	%\input{intro.tex} % A short introduction to LaTeX.
	
	\backmatter
	
	% Use an optional list of figures.
	%\listoffigures % Starred version, i.e., \listoffigures*, removes the toc entry.
	
	% Use an optional list of tables.
	%\listoftables % Starred version, i.e., \listoftables*, removes the toc entry.
	
	% Use an optional list of alogrithms.
	%\listofalgorithms
	%\addcontentsline{toc}{chapter}{List of Algorithms}
	
	% Add an index.
	\printindex
	
	% Add a glossary.
	% \printglossaries
	
	% Add a bibliography.
	\bibliographystyle{plain}
	\bibliography{thesis}
	
\end{document}





%\hypersetup{linkcolor=black}
%\tableofcontents
%\hypersetup{linkcolor=DeepBlue}
%\newpage

%\begingroup
%\let\clearpage\relax
%\endgroup

\nocite{*}



\end{document}
